\vfill World of Warcraft's New Way to Get You Hooked

### 2013-04-01 paper.tex
Cleared out the excess for redemption.

%%%%%%%%%%%%%%%%%%%%
%%% Known Facts  %%%
%%%%%%%%%%%%%%%%%%%%

\section{Facts}

Blizzard Entertainment's World of Warcraft is an incredibly popular game with over 10 million subscribers. \cite{WoWPlayerCount} Players of World of Warcraft spend an average of 22 hours per week playing. \cite{PlayerMotivations} World of Warcraft is free to play provided the player is willing to accept some restrictions. None of those restrictions hinder the number of hours one can play. \cite{StarterEditionFAQ} Multiple studies have shown that time spent playing a MMORPG, like World of Warcraft, correlates to addiction level in some individuals. \cite{ProblemUsageAmongHighlyEngagedPlayers}\cite{JustAComputerGameRight} The intensity with which many players play World of Warcraft or other MMORPGs has led to multiple deaths. \cite{SnowlyDeath}\cite{ShawnDeath}\cite{ParentsBlameWoWForDeath}\cite{DeathAfter50Hours}\cite{YouStoleMySwordNowYouDie}

%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Research Question %%%
%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Extant Arguments from External Sources %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Arguments}
\subsection{Arguments Affirmative}
\subsubsection{Differing Player Archetypes}
Users play MMORPGs for a variety of reasons. Sweeping generalizations are often made without basis by grouping all players into a single type that is prone to addiction. \cite{PlayerMotivations}
\subsubsection{More Time Online Does Not Mean Addiction}
Many people who are characterized as spending ``excessive'' (not defined in the article) time online do not show symptoms of addiction. These individuals show signs of psychological fulfillment. Often these people would be unable to find such fulfillment elsewhere due to disabilities, circumstances, or social discomfort. Some people who are isolated or autistic find comfort in the use of the internet and games to create social relationships they would not have cultivated otherwise. \cite{IsThereEvidenceOfInternetAddiction}
\subsubsection{Excessive Users Do Not Exhibit Addiction Symptoms}
Users that spend more time than average playing the game exhibit social fulfillment. Many users who play more than average are playing MMORPGs for social aspects of the in-game world. These users would just as likely find fun elsewhere. \cite{ExcessiveUseForSocialAspects}
\subsection{Arguments Negative}
\subsubsection{More Intense Play By A Younger Audience Leads to Addiction}
``MMORPGs are played more intensely than [offline] video games,'' and are played primarily by a younger audience. Persons under the age of 27, the majority of players of MMORPGs, are far more susceptible to addiction due to a lack of self identity and self confidence. \cite{YoungerAudiencesGetAddicted}
\subsubsection{Increased Time Playing Leads to Life Problems}
The longer a person plays World of Warcraft, the more likely they are to experience issues in their lives. \cite{ProblemUsageAmongHighlyEngagedPlayers}
\subsubsection{Addictive Games Have a Social Responsibility}
Current efforts to warn players of online games' addictiveness are minimal and often non-existent. Warning labels outlining the possibility of addition should be printed on game boxes and companies should contact users who play in excess of a set threshold. \cite{MMORPGSocialResponsibility}\\

Arguments that World of Warcraft's model \emph{is} ethical boil down to the fact that people are complex and a simple correlation of time spent playing to a person being psychologically addicted is too far of a stretch. People play for a variety of reasons, many of which are psychologically fulfilling. Arguments that propose that World of Warcraft is causing addiction bring up that often players have immature brains and a lack of social identity due to age, leading to increased risk of addiction. They also cite studies that show increased play times leading to life issues and that the companies have social responsibility to at least warn consumers of possible psychological consequences.

@@

\subsection{SE Code Section 2.07\\Identify, document, and report significant issues of social concern, of which they are aware, in software or related documents, to the employer or the client.}
Software Engineers are required to make note of ``significant issues'' and make sure that users are aware of these issues. It's no stretch to consider psychological addiction to be a ``significant issue of social concern'' and therefore it is Blizzard Entertainment's responsibility to ``identify, document, and report'' this issue. As before, ``the client'' is the player and the ``software'' is World of Warcraft.

During a panel at D.I.C.E. 2012 \cite{DICEInterview} Frank Pearce answered a question about World of Warcraft addiction:
\begin{quote}
Mike Epps: It looks like I've put a good half man-year into one of my [World of Warcraft] characters. I'm not gonna say ``How do you sleep at night?'' ... But, do you think about that? The implications of this industry, of entertaining?\\
\\
Frank Pearce: Yeah, I mean we think about it all the time. ... If there's something that you're really really passionate about and you're consuming the media, whether it's videogames or a book or a television series... that means someone's created some great content, but I mean we have to consume it in moderation.
\end{quote}
Frank Pearce, and surely others at Blizzard, are aware of the issue of addiction, taking care of the clause ``of which they are aware.''

The code section in this case becomes ``World of Warcraft's software engineers shall identify, document, and report the issue of addiction to World of Warcraft to its players.''

World of Warcraft has a number of loading screens: when you log into the game, move across the map, and do a variety of other things.\footnote{Having over two months of play time in-game with the last play session less than a week ago, I consider myself an expert in the field of World of Warcraft, and thus needn't cite anyone but myself.} On each of these loading screens is a little segment of text with various ``tips'' that provide humor, help you in the game, or give you some advice. One such tip is: ``Tip: Remember to take all things in moderation (even World of Warcraft!)'' \cite{LoadingScreenTips} Blizzard's software engineers are specifically identifying and documenting the issue of addiction and suggesting time away from World of Warcraft.

In World of Warcraft one of the primary goals is to level up your character, unlocking new skills and becoming more powerful. In order to level up you complete quests and kill enemies, both of which give you experience points. When you collect enough experience points, you level up. If you log out of the game when standing in certain locations, usually at ``inns'' in big towns, you become ``rested'' for some period of time which increases the longer you are logged out. Being rested when you log back in, you gain twice as much experience from killing enemies as when you're not rested. This state of being ``rested'' wears off as you play and adventure in the world. In essence, logging out of the game actually helps you in the game.

Some experts have suggested warning labels for games like World of Warcraft as well as contacting players who exhibit signs of addiction, such as play sessions in excess of some period of time. \cite{MMORPGSocialResponsibility} I see nothing wrong with a few added measures that go hand in hand with the idea of identifying and reporting the issue of addiction.

Blizzard is clearly making the effort to get users to play in moderation. They even offer rewards for being offline. They are not using warning labels or contacting players, however.

\subsubsection{Ethical Conclusion}
Blizzard Entertainment and its software engineers have shown knowledge of the addiction issue and have identified and reported it in-game as well as implemented rewards that encourage off-line time.

Although Blizzard could be doing more, they have put in effort to help users and therefore they are ethical in identifying, documenting, and reporting the issue of addiction as well as going beyond the requirements and encouraging time away from the game.

\section{Conclusion}
There is no link between time spent doing an activity and addiction to it, for every person the timeline is different.

Blizzard Entertainment has moderated its interests, and documented the possibility of addiction. Blizzard has also gone above and beyond and improved the public's welfare while offering players rewards for \emph{not} playing.

World of Warcraft is a game that pushes its users to interact with each other and offers users the chance for achievement, socializing, and immersion in a world unlike their own. Through its mechanics and gameplay users learn valuable life and leadership skills and can interact with others to achieve a common goal.

Is Blizzard's use of World of Warcraft: Starter Edition ethical? Yes.

%end the two column format
\end{multicols}
\newpage

%cite all the references from the bibtex you haven't explicitly cited
\nocite{*}

>>>>

@@

@@

\end{multicols}
\newpage
